Classless Java

### 2016-03-21 OOPSLA16/sections/Implementation.tex
Section 3: flow chart, screenshot and text.

\subsection{Compilation Agents}
Java supports compilation agents\cite{compilationagents}.
That is, Java libraries can interact with the Java compilation process,
acting as a man in the middle between the
generation of the abstract syntax tree and the Java Bytecode.
\marco{is this correct?}\bruno{my impression is that Lombok happens
  \emph{before} the AST is compiled (not between AST and bytecode generation)}
This process is facilitated by frameworks like Lombok~\cite{lombok}:
a Java library that aims at removing (or
reducing) Java boilerplate code via
annotations. Classless Java was created using Lombok.
Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
flow of \mixin annotation processing in Lombok.\bruno{Haoyuan, please
  explain the diagram.}

@@

\begin{figure}[t]
\centering
\includegraphics[width=3in]{pdfs/lombok2.png}
\caption{The flow chart of Lombok annotation processing. 
\marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
but there is also the "javac" path, and here we call the graph "Lombok annotation processing".
May be this should be "obj" processing?
}}\bruno{Haoyuan, please remove references to Eclipse in the Figure,
that way it works both for Eclipse and Javac}
\label{fig:lombok}
\end{figure}

@@

\paragraph{Tool support:}
Features written in Lombok integrate and are supported directly in the
language, and are often also supported by (most) tools.  For example in Eclipse, the processing is
performed transparently and the information of the interface from
compilation is captured in the ``Outline'' window.\bruno{Haoyuan: show
and explain figure here.} This includes all
the methods inside the interface as well as the generated ones.  This
also means that the IDE functionality for content assist and
autocomplete will work for the newly generated methods.

>>>>

\subsection{Compilation Agents}
Java supports compilation agents\cite{compilationagents}.
That is, Java libraries can interact with the Java compilation process,
acting as a man in the middle between the
generation of the abstract syntax tree and the Java Bytecode.
\marco{is this correct?}\bruno{my impression is that Lombok happens
  \emph{before} the AST is compiled (not between AST and bytecode generation)}
This process is facilitated by frameworks like Lombok~\cite{lombok}:
a Java library that aims at removing (or
reducing) Java boilerplate code via
annotations. Classless Java was created using Lombok.
Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
flow of \mixin annotation processing in Lombok.\bruno{Haoyuan, please
  explain the diagram.}
\haoyuan{Source code in a Java program is first parsed into its syntactic structure,
which we call abstract syntax tree (AST). The AST is passed to Lombok, and a corresponding
AST defined in Lombok is generated for the annotated node. An annotation like \mixin is declared
inside Lombok, together with an implementation of its (Eclipse or Javac) handler. The handler is
free to modify the information of the annotated node, or even inject new nodes (like methods, inner classes,
etc). Finally, Java compiler works on the modified AST to generate bytecode. During the compilation,
no source code is changed, and no new files are created, bytecode takes the only difference.}

@@

\begin{figure}[t]
\centering
\includegraphics[width=3in]{pdfs/lombok3.png}
\caption{The flow chart of \mixin annotation processing.
\marco{I'm not sure this helps: It is trying to show the "eclipse" path, that is the one we implemented,
but there is also the "javac" path, and here we call the graph "Lombok annotation processing".
May be this should be "obj" processing?
}}\bruno{Haoyuan, please remove references to Eclipse in the Figure,
that way it works both for Eclipse and Javac}\haoyuan{done}
\label{fig:lombok}
\end{figure}

@@

\paragraph{Tool support:}
Features written in Lombok integrate and are supported directly in the
language, and are often also supported by (most) tools.  For example in Eclipse, the processing is
performed transparently and the information of the interface from
compilation is captured in the ``Outline'' window.\bruno{Haoyuan: show
and explain figure here.} This includes all
the methods inside the interface as well as the generated ones. \haoyuan{In Figure~\ref{fig:screenshot} (wrong),
the screenshot presents the Outline window of Eclipse, where the generated methods including \lstinline{Point2D.of}, \lstinline{Point3D.of},
\lstinline{Point3D.withX} and \lstinline{Point3D.withY} are visible to users, showing their types and modifiers.} 
This also means that the IDE functionality for content assist and
autocomplete will work for the newly generated methods.

\begin{figure}[h]\label{fig:screenshot}
\centering
\includegraphics[width=3.3in]{pdfs/screenshot2.png}
\caption{Generated methods shown in the Outline window of Eclipse.}
\end{figure}
